Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-11-13 paper/appendix.tex
wip

Three distinct thresholding rules can be used to binarize $\text{diag(\hat H)}$ and tbus explicitly recover causal features under noise. (In the MEG experiment, we implement the third one).

First, given known signal-to-noise ratio, one can derive the bound above which a feature should considered to be causal. Indeed, Equation $\ref:diagk$ implies that the $k$ first diagonal elements of $\hat H$ are bounded:  % we assume that $F$ has full rank over $Img(E)$.

@@

The average value $\mu$ of non-zero coefficients of $\text{diag}(\hat H)$ is the trace of $\hat H$ divided by $k$. In practive $\mu$ can be estimated as :
\begin{equation}
\mu = \frac{Var(X)}{Var(X)+Var(N)}
\end{equation}

The decision threshold between causal and non-causal elements is thus a fraction $\mu$. This threshold is to be adjusted depending on the type I and type II risks.

Second, $\text{diag{\hat H}}$ can be binarized with the Sonquist-Morgan criterion \citep{sonquist_morgan}, a clustering procedure separating small and large values. This procedure maximizes the ratio of inter-group variance ands minimizes the intra-group variance, over all possible splits of the diagonal into $p$ largest values and $d_x-p$ smallest values.
%
Letting $m_0$ and $m_1$ be the average values of the two clusters, $p$ and $d_x-p$ their size, and $v$ the total variance of the sample, Sonquist-Morgan criterion maximizes \citep{sonquist_morgan} \citep{kass75}:
\begin{equation}
  \frac{p(d_x-p)}{d_x} \frac{(m_1 - m_0)^2}{v}
\end{equation}
%
Finally, second-order statistics across multiple datasets can be used to identify the elements of $\text{diag}(\hat H)$ that are sgnificantly different from 0. This procedure is detailed in the methods.

>>>>

In case of noise, B2B yields non binary $\hat E$. Three thresholding rules can be used to binarize its values thus explicitly recover causal features.

First, given known signal-to-noise ratio,the threshold above which a feature should considered to be causal can be derived analytically. Indeed, Equation $\ref:diagk$ implies that the $k$ first diagonal elements of $\hat H$ are bounded:  % we assume that $F$ has full rank over $Img(E)$.

@@

The average value $\mu$ of non-zero coefficients of $\text{diag}(\hat H)$ is the trace of $\hat H$ divided by $k$, and can be computed as
\begin{equation}
\mu = \frac{Var(X)}{Var(X)+Var(N)}
\end{equation}

The decision threshold between causal and non-causal elements is thus a fraction $\mu$, whose proportion arbitrarily depends on the necessity to favor type I and type II errors. In practice, we cannot use this procedure for our MEG study, because signal-to-noise ratio is unknown.

Second, $\text{diag{\hat H}}$ can be binarized with the Sonquist-Morgan criterion \citep{sonquist_morgan}, a non-parametric clustering procedure separating small and large values in a given set. This procedure maximizes the ratio of inter-group variance while minimizing the intra-group variance, over all possible splits of the diagonal into $p$ largest values and $d_x-p$ smallest values. Let $m_0$ and $m_1$ be the average values of the two clusters, $p$ and $d_x-p$ their size, and $v$ the total variance of the sample, Sonquist-Morgan criterion maximizes \citep{kass75}:
\begin{equation}
  \frac{p(d_x-p)}{d_x} \frac{(m_1 - m_0)^2}{v}
\end{equation}

This procedure assumes that there exists at least one causal and at least one non-causal feature.
%
Third, second-order statistics across multiple datasets can be used to identify the elements of $\text{diag}(\hat H)$ that are sgnificantly different from 0. This procedure is detailed in the method section of our MEG experiment.

Overall, these three procedure thus varies in their additional assumptions: i.e. (1) a known signal-to-noise ratio, (2) the existence of both causal and non-causal factors or (3) independent repetitions of the experiment.
